---
title: "Project 4 - Document Classification"
author: "Nicholas Kunze"
date: "2024-04-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project I will be using the [Naive](https://learning.oreilly.com/library/view/doing-data-science/9781449363871/ch04.html#idm46107867092920) [Bayes](https://www.ibm.com/topics/naive-bayes) classifier machine learning algorithm to make predictions on whether an email is spam or not. IBM, the company I work at, has a lot of education around ML algorithms and I leveraged that here in the project. I've also previously worked with a team on something similar (not Naive Bayes) using a TensorFlow model to predict credit card fraud, the public version of which is hosted [here](https://github.com/IBM/ai-on-z-fraud-detection).

First, we'll need to ingest our data and prepare it; preparing it will involve creating the 'metadata' around each email by creating our corpora. The important part here is that the data we are using is already marked as fraudulent, spam, or not, ham. We then split this data into two groups, one for our model training and the other to test how accurate our trained model is. I usually split these into groups of sized 90/10 or 95/5.

As stated, after this we train our model

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
